Procedural generation of levels for the angry birds videogame using evolutionary computation\\

### 2019-05-16 IEEE_ConferenceOnGames/new_article.tex
abstract

\begin{abstract}
In this paper we propose an evolutionary computation-based system to generate
levels for the Angry Birds video game. The levels themselves are composed of a
set of birds that the player can use, a certain amount of pigs that must be
destroyed in order to progress to the next level and a given number of pieces
that conform structures that may or may not protect the pigs from the player.
The focus in this stage of the work is the generation of diverse structures that
can be played in the game having the additional characteristics of being
interesting and playable. We propose a multi-layered search, first constructing
composite structures from basic components, to then build more complex
structures building from these components, following an open-ended evolution
approach in which the evolution of structures is not guided towards a single
objective but is rather free to evolve and generate novelty or diversity. The
objective is then that the proposed levels are evaluated by how complex they
become and how different they are from the rest of the group. The experiments
conducted show that a evolutionary approach allows the generated levels to be
unique since, on most cases, the way they are presented on screen is a way a
normal human would have trouble to create but at the same time are interesting
to see.
% Another revision to English would be good.
\end{abstract}

@@

    \section{Proposed Method}
    \label{prop_method}

    % Is this explanation undertandable ? JSH
    In this section we describe the approach used in this work, explaining first,
    the restrictions of the game, then giving a quick overview of how we create new 
    composites to be used by the algorithm, then explaining how the candidate solutions 
    are represented and how this are modified and evolved by the algorithm, and finally 
    we present our method to evaluate the candidate solutions.

    The game contains a total of 11 block shapes as shown in Figure \ref{piece_list}.
    These basic blocks cannot be modified, it is not possible to add more or modify the existing ones,
    with this in mind the purpose is to create composites of these blocks in
    different locations and angles in order to create structures that can be used as
    building blocks for more complex structures.

@@

    Since there is no way to \textit{add} a new composite to the pool of blocks,
    then a way to combine both is needed, in order to do this first the original
    11 pieces of the game will be converted into composites of a single element.
    A composite is defined as a lists of blocks, then a \textit{composite} of
    one element will be created for each piece. Composites are the first layer
    the building blocks for the structures to come. A composite could include
    different types of blocks, rotated at a certain angle and compose a basic
    structure. An ideal composite could support other composites on top with out
    falling apart, but this is not required. A composite is measured using
    imaginary bounding box, defined by calculating the lengths of all the pieces,
    finding the borders of said pieces and then finding the points that could
    define a box around all of them, then use this defined box to calculate the
    height and width of the new composite and then calculate from the center
    point of said box the offsets from the center of the group to the center of
    each respective piece as shown on figure \ref{bounding_boc_calc}, this
    information is needed in order to add it as a new composite so it can be
    used to place the pieces in their respective new positions when the files
    needed for the simulation are being created. This is similar to creating a
    group of elements in a drawing application.

    %% 
    Good composites can be found using an evolutionary approach proposed earlier 
    by Calle and Merelo, where basic structures are generated by searching a space
    of falling blocks and selecting those that remain standing. This algorithm is 
    not capable of creating complex structures, but it can generate very novel basic
    structures. We can see that the evolution at this level can benefit from having
    more freedom in how basic structures are built. We do not want to form piles
    or sequences of blocks, this will be done at a higher level in the composite hierarchy.

    At this level the composite generator has the following steps:
    \begin{enumerate}
    \item Generate the first generation of the population, consisting of
    randomly generated list of blocks.
    \item Run a Simulation.
    \item Evaluate each individual of the population.
    \item If a member of the population has remaining pieces obtain the location
    and angle of the pieces, assign a fitness.
    \item Do the selection, crossover and mutation operations.
    \item Repeat the cycle.
    \item From the best individuals select composites. Several algorithms could
    be proposed for this step, for instance selecting clusters of blocks.
    \end{enumerate}

    Another option is to just manually generate interesting composites to be used by the algorithm.
    We could say that the previous approach is an option for a procedural generation of composites.

    \begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/list_pieces.png}}
    \caption{List of pieces in the angry birds game.}
    \label{piece_list}
    \end{figure}
    % Explain a bit about the competition, from the introduction
    The Angry Birds level generation track is a competition where the objective is 
    to generate levels that are \textit{fun} and \textit{enjoyable} however they 
    also have to be challenging but not impossible. However the rules for the competition 
    specify that a configuration file must be used to
    generate the levels of the game in a limited amount of time, the information on
    the configuration file is first the number of \textit{different} levels
    that must be generated, the second value is the different materials \textit{not
    allowed} to be used, for instance a group of generated levels must not have
    circles which composition material is ice or stone, moreover it does not have to
    integrate a triangle piece with ice, stone or wood material so this combinations
    must be removed from the pool of options, the third configuration value is the
    \textit{number of pigs} that must be placed in a level, and finally the
    \textit{limit} of time that the system has to create the required amount of
    levels.

    In our original approach, composites could added to the structure generator
    algorithm at runtime as a mutation of an individual. The structure generator
    is explained in the next section.
    In our current approach the composite generator is only
    run once at the beginning, as a setup process.

>>>>

\begin{abstract}
In this paper, we propose an evolutionary algorithm that follows an
open-ended evolution approach to generate levels for the Angry Birds video
game. The levels themselves are composed of a set of birds that the player
can use, a certain amount of pigs that they must destroy, and a given number
of pieces that conform structures that may or may not protect the pigs from
birds. The current goal is the generation of diverse structures playable in
the game, having the additional characteristics of being fun and enjoyable.
We propose a multi-layered search, first constructing composite structures
from basic blocks, to then build more complex structures building from these
composites, following an open-ended evolution approach in which the
evolution of structures is not guided towards a single objective but is
rather free to evolve and generate novelty or diversity. The fitness
function we use to evaluate the proposed levels considers how complex they
have become and how different they are from the rest of the group. The
experiments conducted show that an evolutionary approach allows the
generated levels to be novel and interesting to play.
% Another revision to English would be good.
% Done - Mario
\end{abstract}

@@

    \section{Proposed Method}
    \label{prop_method}  
    % Is this explanation undertandable ? JSH
    In this Section, we describe the approach used in this work, explaining
first, the restrictions of the game, then giving a quick overview of how we
create new composites to be used by the algorithm, then we define how candidate
solutions are to be represented by the algorithm. We then describe our method to
evaluate candidate solutions. Finally, we present details of the evolutionary
algorithm.

The game contains a total of 11 block shapes, shown in Figure
\ref{piece_list}. These basic blocks cannot be modified, and we can not add
more. With this in mind, we propose to create additional elements by creating
composites of these blocks. Atomic blocks can be positioned in different
locations and at certain angles, inside the composite. The resulting composites
are themselves the new building blocks of more complex structures.

@@

Since there is no way to \textit{add} a new composite to the pool of blocks, we
first need a way to combine them. To do this, first, the original 11 blocks of
the game need to be converted into composites of a single element.  We define a
composite as a list of blocks; then  \textit{composite} with a single block,
will be created for each original block. Composites are the first layer, the
building blocks for the structures to come. A composite could include different
types of blocks, rotated at a certain angle and compose a basic structure. An
ideal composite could support other composites on top without falling apart, but
this is not required. A composite is measured using an imaginary bounding box,
defined by calculating the lengths of all the pieces, finding the borders of
said pieces and then finding the points to define a box around all of them. We
then use this defined box to calculate the height and width of the new
composite. Then we derive from the center point of the said box, the offsets for
each block in the composite, as shown in figure \ref{bounding_boc_calc}. We need
this information to treat each composite as if it was a single block; this will
enable the placements of the composites in their respective new position when we
create the files needed for the simulation. This process is similar to creating
a group of elements in a drawing application.

    %% 
Suitable composites can be found using an evolutionary approach proposed
earlier by Calle and Merelo, where basic structures are generated by freely
searching a space of falling blocks and selecting those that remain
standing. This algorithm is not capable of creating complex structures, but
it can produce original basic structures. We can see that evolution at this
level can benefit from having more freedom when building basic structures.
At this time, we do not want to form piles or sequences of blocks, this will
be done later, at a higher level in the composite hierarchy.    
At this level the composite generator has the following steps:
\begin{enumerate}
\item Generate the first generation of the population, consisting of
randomly generated list of blocks.
\item Run a Simulation.
\item Evaluate each individual of the population.
\item If a member of the population has remaining pieces obtain the location
and angle of the pieces, assign a fitness.
\item Do the selection, crossover and mutation operations.
\item Repeat the cycle.
\item From the best individuals select composites. Several algorithms could
be proposed for this step, for instance selecting clusters of blocks.
\end{enumerate}

Another valid option is to generate interesting composites manually.  The
algorithm can later use the composites. We could say that the approach we
explained earlier, is an optional procedural generation of composites.

\begin{figure}[htbp]
\centerline{\includegraphics[width=80mm]{Images/list_pieces.png}}
\caption{List of blocks in the angry birds game.}
\label{piece_list}
\end{figure}
    % Explain a bit about the competition, from the introduction
The Angry Birds level generation track is a competition where the objective
is to generate levels that are \textit{fun} and \textit{enjoyable}. However,
they also have to be challenging but not impossible. The rules for the
competition specify that a configuration file must be used to generate the
levels of the game in a limited amount of time. The file contains several
restrictions; we explain each one next. The number of \textit{different}
levels that must be generated. The materials that are \textit{not allowed}.
For instance, a group of generated levels must not have circles which
composition material is ice or stone. Moreover, a material with a veto,  can
not integrate others. The third configuration value is the \textit{number of
pigs} that we must place in a level. Finally, the \textit{limit} of time
that the system has to generate the required amount of levels.
